---
title: <span style="color:#B7F06C;font-size:60px">Assignment 5.2</span>
output:
  html_document: 
    toc: yes
    toc_float: yes
    toc_depth: 5
    theme: darkly
    highlight: tango
author: |
  <span style="color:grey;font-size:17px">Author: Patrick Weatherford</span>
date: |
  <span style="color:#5c5c5c;font-size:14px">Last modified: `r format(Sys.time(), '%d %B %Y')`</span>
    
---

<!-- CSS Style -->
<style>

  h3 {
    color: #a56cf0;
    font: bold;
    }

  
  h4 {
    color: white;
    }
  
  .tocify-header {
    color: #a56cf0;
    }
    
  
  .list-group-item.active, .list-group-item.active:hover, .list-group-item.active:focus {
    background-color: #4b2b75;
    border-color: #4b2b75;
    }
    
  div.main-container {
    max-width: 2000px
    }
    
  #TOC {
    margin: 25px 20px 20px 0px;
    width: 30%;
    max-width: 300px
    }
    
  .list-group-item {
    background-color: #2b2b2b
    }
  
  .tocify-subheader {
    color: white;
    }
    
  .tocify-subheader .tocify-subheader {
    color: #b0b0b0;
    }
  
  body {
    color: grey;
    background-color: #222222;
    }
  
  
  hr {
    border: 1px solid;
    color: grey;
    }
    
  BLOCKQUOTE {
    font-size: 15px;
    color: #ebc634;
    }
    
  pre:not([class]) {
    background-color: #222222;
    color: #B7F06C;
    border: none;
    }
    
  pre.sourceCode {
    background-color: #454545;
    color: white;
    white-space: pre-line;
    }
    
  pre.numberSource {
    margin-left: 2em;
    }

  /*quotations and string characters*/
  code span.st, code span.sc {
    color: #f5ea78;
    }

  /*numbers | TRUE/FALSE*/
  code span.dv, code span.fl, code span.cn {
    color: #63aaff;
    }

  /*keyword*/
  code span.fu, code span.cf {
    color: #f73e7f;
    }

  /*comment | operators*/
  code span.co, code span.ot {
    color: #8c8c8c;
  }
    
  tbody {
    color: #adadad
    }
    
  thead {
    color: #B7F06C
    }
    
  tr:hover, .table-striped>tbody>tr:nth-of-type(odd):hover {
    background-color: #4b632b
    }
    
  code {
    box-sizing: initial
    }
    
</style>

***
<br>
```{r default-chunk-settings, include=FALSE}
knitr::opts_chunk$set(
  comment=''
)
options(width = 125) 
# matrix(runif(100), ncol = 30)
```


### Load Packages for Assignment
Load appropriate packages for the assignment
```{r, message=FALSE}
library(tidyr)
library(dplyr)
library(rmarkdown)
library(gridExtra)
library(ggplot2)
library(readxl)
library(stringr)
```
<br>


### Read Data Source Into Data Frame
```{r}

housing_df <- readxl::read_excel(path="C:/Users/patwea/OneDrive - Bellevue University/BU/DSC 520 - Statistics for Data Science/Patrick.Weatherford-DSC520/dsc520-clone/data/week-7-housing.xlsx"
                   , sheet="Sheet2")

hdf <- housing_df

```
<br>


### Data Transformation

Using either the same dataset(s) you used in the previous weeks’ exercise or a brand-new dataset of your choosing, perform the following transformations (Remember, anything you learn about the Housing dataset in these two weeks can be used for a later exercise!)

<ol style="list-style-type: lower-alpha">

<li> Using the dplyr package, use the 6 different operations to analyze/transform the data - GroupBy, Summarize, Mutate, Filter, Select, and Arrange – Remember this isn’t just modifying data, you are learning about your data also – so play around and start to understand your dataset in more detail
```{r}

# Calculate outlier criteria
quants <- stats::quantile(hdf$`Sale Price`, probs=c(.25, .75))

q1 <- as.numeric(quants["25%"])
q3 <- as.numeric(quants["75%"])
iqr <- IQR(hdf$`Sale Price`)

low_out <- q1 - (iqr*1.5)
high_out <- q3 + (iqr*1.5)


new_hdf <- hdf %>%
  dplyr::filter(`Sale Price` > low_out & `Sale Price` < high_out) %>%
  dplyr::group_by(current_zoning) %>%
  dplyr::summarise("Zone Median Price"=median(`Sale Price`)) %>%
  dplyr::right_join(hdf, by="current_zoning") %>%
  dplyr::mutate("Above Median Price?"=ifelse(`Sale Price`>`Zone Median Price`,"Y","N")) %>%
  dplyr::select("Sale Date", "Sale Price", "Zone"="current_zoning", "Zone Median Price", "Above Median Price?") %>%
  dplyr::arrange(desc(`Zone Median Price`), `Sale Price`) %>%
  rmarkdown::paged_table(options = list(rows.print=30))

new_hdf

```
<br>


<li>Using the purrr package – perform 2 functions on your dataset.  You could use zip_n, keep, discard, compact, etc.
```{r}

hdf[,c("Sale Price", "square_feet_total_living", "sq_ft_lot")] %>% 
  purrr::map(sd)

hdf$sale_warning %>%
  purrr::keep(function(x) !is.na(x)==TRUE) %>% 
  data.frame() %>% 
  rmarkdown::paged_table(options = list(rows.print = 20))

```
<br>


<li>Use the cbind and rbind function on your dataset
```{r}

id <- c(1,2,3,4,5)
char_var <- c("a","b","c","d","e")

my_df <- data.frame(id, char_var)

cap_char_var <- c("A","B","C","D","E")

new_ob <- c(6,"f","F")

my_df <- cbind(my_df, cap_char_var)

my_df <- rbind(my_df, new_ob)
  
my_df
```
<br>


<li>Split a string, then concatenate the results back together
```{r}

str_df <- data.frame("str"=c("yo-123", "blahblah-3653", "HakunaPatata-777"))

str_df2 <- str_df$str %>% 
  stringr::str_split(pattern = "-") %>% 
  base::Reduce(rbind, .) %>% 
  base::data.frame() %>% 
  dplyr::mutate(combined_char = paste0(.$X1, "-", .$X2))
  
str_df2


```
<br>

